Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-10-05 paper.tex
finish edits from dad

Throughout most of the shortage, the aggregate outflow of Pu from separated
inventory is roughly the same in cases MF and QF.  However, due to the longer
time steps in case QF, the shortage starts slightly sooner because Pu inventory
must be maintained at a higher limit as described above.  The small difference
in outage start time has feed-back effects.  Not only are extra reactors
offline, but those extra reactors are also discharging less aggregate spent
fuel for recycling.  This farther increases the discrepency between the two
cases. Eventually the shortage in case MF ends slightly before it does in case
QF. This feedback effect occurs because the reactors in question have a
conversion ratio greater than 1.0.

@@

The in-flows shown for cases MF and QF show a periodic pulsing that begins
during the shortage in about year 120.  This pusling actually has nothing to
do with the shortage and is caused by the retirement of fast reactors.  The
first fast reactors are deployed in year 35 and have an 80 year lifetime.
When they retire, they discharge an entire core's worth of fuel rather than a
single batch.  After a 7-year cooling period, the first discharged full cores
begin to make it through the recycle loop in year 122.  Another interesting
feature visible in all four cases in Figure \ref{fig:puinv} is the minority
fraction of out-flow points that are higher than the bulk majority. These
higher out-flows occur once every 21 months when new reactors are built. The
new reactors require a full core of fresh fuel rather than the single batch
needed when just refueling.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv.eps}
    \caption[Separated Pu inventory and flow]{
        Separated Pu inventory and flows for all four cases.  In-flows for
        cases MI and QI are omitted because they are noisy and obscure other
        useful information.  A larger time step results in larger per time
        step material flows. These larger flows mean more material is not
        available for supplying to requesters.  This information lag causes
        more supply constraints than occur with smaller time steps -
        illustrating the \emph{inventory drawdown effect}. Another result of
        this effect is the larger standing Pu inventory during the shortage
        with larger time steps that can be seen in case QF than in case MF.
    }
    \label{fig:puinv}
\end{figure*}

Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated plutonium
inventory during the shortage generally indicates fewer unfueled reactors.
Lower inventory levels mean available Pu is being utilized more efficiently
rather than sitting idle.  Cases QI and QF with their longer time steps show
higher inventories indicating a more severe shortage.  Post shortage, they
also fall a bit behind on building up Pu stocks due to the delayed
contributions of more unfueled reactors to the separated Pu pool.  The
reactors in all the scenarios request fresh fuel only when they need it. While
this is an unrealistic behavior for real-world facilities, it is closer to a
globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
would also have the same effect of increasing idling plutonium quantities and
would make the fuel shortage worse overall.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare.eps}
    \caption{
        Separated Pu inventories for all four cases.  Differences in modeling
        choices are not particularly significant until about year 100 when
        fuel shortages begin.  These differences gradually decay away starting
        in year 140 after the fuel shortages have ebbed.  Higher aggregate Pu
        inventory  during the shortage is correlated with a longer during and
        post shortage recovery period.
    }
    \label{fig:puinv-compare}
\end{figure*}

@@

Different modeling choices can have a significant impact on the outcome of a
simulation.  With discrete reactor modeling many factors must be considered in
order to ensure the integrity of conclusions.  Individual reactor outage
modeling can be very useful for certain analyses, but it can reduce result
quality if things such as cycle staggering are not handled appropriately.  The
fleet and discreet models used here are just two points in a multi-dimensional
spectrum of modeling choices. If fuel sharing is of interest, but cycle
staggering is not, individual reactors can be used with the power capacity
lowered and the outage reduced to zero duration making the capacity factor
included in the reactor's power capacity.

@@

Increasing time step duration increases facilities' standing inventory
requirement during supply constrained periods.  Facilities can provide no more
inventory than they can hold on a single time step.  So increasing time step
duration for a facility with unconstrained supply generally means higher
standing inventories are required in order to achieve equal throughput.
Facilities with large inventory buffer that are unconstrained are not affected
significantly by the time step duration when operating in isolation.  However,
even when facilities do not have individual inventory or throughput limits,
the time step can still have a significant effect. In recycle loops there is a
constraint on how fast material is introduced through the loop (i.e. the
discharge rate of spent fuel from reactors).  The \emph{inventory drawdown
effect} can impact the aggregate throughput of the recycling loop by affecting
minimum bounds on idling material quantities.

>>>>

Throughout most of the shortage, the aggregate outflow of Pu from separated
inventory is roughly the same in cases MF and QF.  However, due to the longer
time steps in case QF, the shortage starts slightly sooner because Pu inventory
must be maintained at a higher limit as described above.  The small difference
in outage start time has feed-back effects.  Not only are extra reactors
offline, but those extra reactors are also discharging less aggregate spent
fuel for recycling.  This increases the discrepency between the two
cases. Eventually the shortage in case MF ends slightly before it does in case
QF. This feedback effect occurs because the reactors have a
conversion ratio greater than 1.0.

@@

The in-flows shown for cases MF and QF show a periodic pulsing that begins
during the shortage in about year 120.  This pusling actually has nothing to
do with the shortage and is caused by the retirement of fast reactors.  The
first fast reactors are deployed in year 35 and have an 80 year lifetime.
When they retire, they discharge an entire core's worth of fuel rather than a
single batch.  After a 7-year cooling period, the first discharged full cores
begin to make it through the recycle loop in year 122.  Another interesting
feature visible in all four cases in Figure \ref{fig:puinv} is the minority
fraction of out-flow points that are higher than the bulk majority. These
higher out-flows occur once every 21 months when new reactors are built because
new reactors require a full core of fresh fuel rather than the single batch
needed when just refueling.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv.eps}
    \caption[Separated Pu inventory and flow]{
        Separated Pu inventory and flows for all four cases.  In-flows for
        cases MI and QI are omitted because they are noisy and obscure other
        useful information.  When out-flows meet inventory, fuel shortages are taking place. A larger time step results in larger per time
        step material flows. These larger flows mean more material is not
        available for supplying to requesters.  This information lag causes
        more supply constraints than occur with smaller time steps -
        illustrating the \emph{inventory drawdown effect}. Another consequence of
        this effect is the larger standing Pu inventory during the shortage
        with larger time steps that can be seen in case QF than in case MF.
    }
    \label{fig:puinv}
\end{figure*}

Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated plutonium
inventory during the shortage generally indicates fewer unfueled reactors.
Lower inventory levels mean available Pu is being utilized more efficiently
rather than sitting idle.  Cases QI and QF with their longer time steps show
higher inventories indicating a more severe shortage.  Post shortage, cases QI and QF
also fall a bit behind with respect to building up Pu stocks due to the delayed
contributions of more unfueled reactors to the separated Pu pool.  The
reactors in all the scenarios request fresh fuel only when they need it. While
this is an unrealistic behavior for real-world facilities, it is closer to a
globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
would also have the same effect of increasing idling plutonium quantities and
would make the fuel shortage worse overall.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare.eps}
    \caption{
        Separated Pu inventories for all four cases.  Differences in modeling
        choices are not particularly significant until about year 100 when
        fuel shortages begin.  These differences gradually decay away starting
        in year 140 after fuel shortages have ebbed.  Higher aggregate Pu
        inventory  during the shortage is correlated with both a longer shortage and
        a longer post-shortage recovery period.
    }
    \label{fig:puinv-compare}
\end{figure*}

@@

Different modeling choices can have a significant impact on the outcome of a
simulation.  With discrete reactor modeling many factors must be considered in
order to ensure the integrity of conclusions.  Individual reactor outage
modeling can be very useful for certain analyses, but it can reduce result
quality if things such as cycle staggering are not handled appropriately.  The
fleet and discrete models used here are just two points in a multi-dimensional
spectrum of modeling choices. For example, if fuel sharing is of interest but cycle
staggering is not, individual reactors could be used with the power capacity
lowered and the outage reduced to zero duration, including the capacity factor
in the reactor's power capacity.

@@

Increasing time step duration increases facilities' standing inventory
requirement during supply constrained periods.  Facilities can provide no more
inventory than they can hold on a single time step.  So increasing time step
duration for a facility with unconstrained supply generally means higher
standing inventories are required in order to achieve equal throughput.
Facilities with large inventory buffer that are unconstrained are not affected
significantly by the time step duration when operating in isolation.  However,
even when facilities do not have individual inventory or throughput limits,
the time step can still have a significant effect. The \emph{inventory drawdown
effect} can impact the aggregate throughput of supply-constrained recycle loops by affecting
minimum bounds on idling material quantities.
